%This is where the body of your abstract goes, limited to 150 words
%for a thesis, and 350 words for a dissertation or document.  The
%word count limits apply to the regular Abstract in the thesis and
%to the separate Special Abstract.  Use the same text for both; just
%adjust the margins and heading.  The abstract should summarize your
%work.  The UMI booklet listed in the resources section of the U of
%A manual provides some writing tips.  The abstract for a dissertation
%or document may be longer than one page; word count is more important
%than page length in this section.

%If you are doing a paper submission, submit one copy of the special 
%abstract, and two extra copies of your title page, in the box with 
%the final copies of your thesis.  If you are doing an electronic
%submission, you can ignore the special abstract.

Magnetism is a rich subject touching all aspects of physics. My goal with this dissertation is to explore spin and magnetic moments in \emph{relativistic} mechanics from both a quantum and classical perspective. We emphasize the special case of gyromagnetic ratio $g\!=\!2$ and its relationship to the algebraic spin structure of wave equations.

In relativistic quantum mechanics, we investigate generalizations of the Dirac equation for arbitrary magnetic moments for fermions. We analyze the homogeneous magnetic field case and the Coulomb problem for hydrogen-like atoms with emphasis on the role of the anomalous magnetic moment (AMM). We explore alternative approaches which combine mass and the magnetic moment. Classically, we propose a relativistic covariant model of the Stern-Gerlach force via the introduction of a magnetic four-potential. This model modifies the covariant torque equations and unites the Amp{\`e}rian and Gilbertian models for magnetic moments.

We further study (transition) magnetic dipoles in Majorana neutrinos specifically analyzing the relationship between flavor mixing and electromagnetic (EM) fields. We demonstrate EM flavor mixing explicitly in the 2-flavor model and develop a dynamical mass basis with an EM rotation matrix. EM induced neutrino mass splitting is compared to neutrino mass hierarchy.

An interesting application of these theoretical developments is to study primordial magnetization in the early universe during the hot dense electron-positron plasma epoch. We propose a model of magnetic thermal matter-antimatter plasmas. We analyze the paramagnetic characteristics of electron-positron plasma when exposed to an external primordial field. We determine the magnitude of a small polarization asymmetry is sufficient to generate field strengths in agreement with those measured today in deep intergalactic space.

Future research outlooks include: Second order equations for anomalous quantum chromodynamic (QCD) moments, neutrino CP violation in strong EM fields, and fifth-dimension spin dynamics in Kaluza-Klein theory.
