%add acknowledgements here

In my long journey to obtain my PhD, I have benefited from the immense support and friendship of many people in my life. To them I am incredibly grateful.

I first would like to thank Professor Johann Rafelski who as my PhD advisor guided me on my physics journey towards ever increasing heights of knowledge and understanding. Most importantly, I was imbued with a humble awareness of our ignorance towards much of the universe's inner workings. I dearly hope we continue to chip away at those mysteries together. I also would like to thank Victoria Grossack for hosting many students including myself at her and Dr. Rafelski's lovely home. I promise future gluten-free pies which I am now bound to by the publication of this dissertation.

To my fellow research collaborators and simultaneous dear friends, I thank for our wonderful time spent together (in alphabetical order): Stefan Evens, Martin Formanek, Christopher Grayson, Will Price, Cheng Tao Yang. Our group trip to Budapest, Hungary and Prague, Czech Republic this past summer is a memory I will cherish forever. Stefan, I appreciate your eternal enthusiasm and love of older architecture---I owe you a visit, hold me to it! Martin, I won't forget your generous hospitality and your words that ``physics \emph{is} magic.'' Chris, thank you for the kind help the evening before my presentation at PP2023. Thank you Hannah Isaac and Will for saving me from being lost forever within Frankfurt Central station. Cheng Tao, our work on cosmology together has reinvigorated me. I look forward to many more years and publications together.

I thank my closest friends who I've known over many years of fellowship (in alphabetical order): Michael Burrell, Jordan Cohen, Elizabeth Fitzpatrick, Kyle Fitzpatrick, Michael France, Jesse Griggs, Jose Gross, Pedro Gross, Kevin Huffman, Alyssa Playford, and Sagar Samtani. To those in this group who I ride motorcycles with, it's with you I've known true freedom.

I lastly, though equally, thank my family who have supported and been there for me: My mom Maria Steinmetz, my cousins Cindy and Frank Ramos, my aunt and uncle Danuta and Bill Kemp.

Good health and long life to all.